---
title: "R Notebook-Assignment 2"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.

```{r}
# setwd("C:/Users/sputh/Documents/UMD/Spring 20 Courses/758N_DataMining/Lab2") #Don't forget to set your working directory before you start!

library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
library("lubridate")

```

### Loading the data into a data frame

```{r}

dfbOrg <-
  read_csv("data/assignment2BikeShare.csv") %>%
  rename_all(tolower)

```

### Explore the dataset
```{r}
head(dfbOrg)
```

```{r}

skim(dfbOrg)
```

## 1) 
### a) Create the additional variables
#### i) Create a "count" variable
```{r}
dfbOrg <-
  dfbOrg %>%
  mutate(count = registered + casual)

head(dfbOrg)
```

#### ii) Create a month variable
```{r}

dfbOrg <-
  dfbOrg %>%
  mutate(month = months(date))

head(dfbOrg)
```

### b) Scale the data
```{r}

dfbStd <-
  dfbOrg %>% 
  mutate(temp = scale(temp), atemp = scale(atemp), humidity = scale(humidity), windspeed = scale(windspeed))

head(dfbStd)
```

## 2) Basic Regression in R

```{r}

fitAll <- lm(formula = count ~ ., data = dfbStd)
summary(fitAll)
```

## 3) Working with data and exploratory analysis
### a) Add a variable badWeather

```{r}

dfbOrg <-
  dfbOrg %>% 
   mutate(badWeather = ifelse( weathersit == 3 | weathersit ==4, "Yes", "No"))

head(dfbOrg)
```

### b) Scatterplot of count vs atemp

```{r}
dfbOrg %>% 
 ggplot(aes(x=atemp,y=count, color = badWeather)) + geom_point()
```

### c) Scatterplots for casual vs atemp and registered vs atemp

```{r}
# feels like temperature vs casual rides count
dfbOrg %>% 
 ggplot(aes(x=atemp,y=casual, color = badWeather)) + geom_point()

```

```{r}
# feels like temperature vs registered rides
dfbOrg %>% 
 ggplot(aes(x=atemp,y=registered, color = badWeather)) + geom_point()
```

#### iv) Scatterplot for count vs atemo with a smooth line showing trend

```{r}
dfbOrg %>% 
 ggplot(aes(x=atemp,y=count)) + geom_point() + geom_smooth()

```

## 4) More Linear Regression

```{r}

fitOrg <- lm(formula = count ~ month+weekday+badWeather+temp+atemp+humidity, data = dfbOrg)
summary(fitOrg)
```

## 5) Regression diagnostics

```{r}

plot(fitOrg)

```


```{r}
# Check multicollinearirty
car::vif(fitOrg)

```

```{r}
# check autocorrelation
dfbOrg %>% 
  modelr::add_residuals(fitOrg) %>%
  ggplot(aes(date, resid)) +
  geom_line()
```

```{r}
# Revised model with fixes for multicollinearity and autocorrelation
fitOrg <- lm(formula = count ~ date+month+badWeather+temp+humidity, data = dfbOrg)
summary(fitOrg)

```

```{r}
plot(fitOrg)
```

```{r}

# Check multicollinearirty
car::vif(fitOrg)

```

```{r}
# Check autocollinearity
dfbOrg %>% 
  modelr::add_residuals(fitOrg) %>%
  ggplot(aes(date, resid)) +
  geom_line()
```


## 6) Even more regression

```{r}

fitBadWeather <- lm(formula = count ~ badWeather, data = dfbOrg)
summary(fitBadWeather)

```
### c) Include weekday and interaction between badWeather and weekday

```{r}

fitBadWeatherWeekday <- lm(formula = count ~ badWeather*weekday, data = dfbOrg)
summary(fitBadWeatherWeekday)

```

## 7) Predictive Analytics

### a,b) Set the seed and split the data into training and testing sets

```{r}
set.seed(333)

dfbTrain <- dfbOrg %>%
  sample_frac(size=0.8)

dfbTest <-
  dplyr::setdiff(dfbOrg, dfbTrain)

```


### c)
#### i) Model 1
```{r}

fitOrg <- lm(formula = count ~ date+month+badWeather+temp+humidity, data = dfbTrain)
summary(fitOrg)

```

```{r}
resultsOrg <-
  dfbTest %>% 
  mutate(predictedCount = predict(fitOrg, dfbTest))

resultsOrg
```

```{r}

performance <- metric_set(rmse, mae)
performance(resultsOrg, truth = count, estimate = predictedCount)

```

#### i) Model 2

```{r}
fitNew <- lm(formula = count ~ date+month+badWeather+temp+humidity+windspeed, data = dfbTrain)
summary(fitNew)
```

```{r}

resultsNew <-
  dfbTest %>% 
  mutate(predictedCount = predict(fitNew, dfbTest))

resultsNew
```

```{r}

performance <- metric_set(rmse, mae)
performance(resultsNew, truth = count, estimate = predictedCount)

```

```{r}
# Anova test to compare models
anova(fitOrg, fitNew)
```

## 8) More predictive analytics

```{r}

# Split the data in year 2011 as training and 2012 as testing data.
dfbTrain <- dfbOrg %>%
  filter( year(date) == 2011)

dfbTest <-
  dplyr::setdiff(dfbOrg, dfbTrain)
```

```{r}

fitOrg2011 <- lm(formula = count ~ date+month+badWeather+temp+humidity, data = dfbTrain)
summary(fitOrg2011)
```

```{r}

resultsOrg2011 <-
  dfbTest %>% 
  mutate(predictedCount = predict(fitOrg2011, dfbTest))

resultsOrg2011
```

```{r}

performance <- metric_set(rmse, mae)
performance(resultsOrg2011, truth = count, estimate = predictedCount)

```

```{r}
# Split the data in year 2011 as training and 2012 as testing data.
dfbTrain <- dfbOrg %>%
  filter( year(date) == 2011 | (year(date) == 2012 & month(date) <=6))

dfbTest <-
  dplyr::setdiff(dfbOrg, dfbTrain)
```

```{r}
fitOrg18months <- lm(formula = count ~ date+month+badWeather+temp+humidity, data = dfbTrain)
summary(fitOrg18months)
```

```{r}
resultsOrg18months <-
  dfbTest %>% 
  mutate(predictedCount = predict(fitOrg18months, dfbTest))

resultsOrg18months
```

```{r}

performance <- metric_set(rmse, mae)
performance(resultsOrg18months, truth = count, estimate = predictedCount)
```

