---
title: "Assignment 4 R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.

# **Assignment 4**

```{r}
# Load required libraries
library(tidyverse)
library(fpp3)
library(plotly)
library(skimr)
library(lubridate)
```

# Part I
## 1) Data Processing

### a) Load lending club data set

```{r}

tsLCOrg <-
  read_csv("data/lendingClub.csv") %>%  
  rename_all(tolower)

```

### b) Convert dataset to tibble

```{r}
tsLCOrg <- as_tsibble(tsLCOrg,key = state, index = date)
head(tsLCOrg)
```

### c) Inspect and describe the data

```{r}
skim(as_tibble(tsLCOrg))
```


```{r}
str(tsLCOrg)
```

### c) Load NY Economy indicators dataset
```{r}
nyEconInd <-
  read_csv("data/nyEcon.csv") %>%  
  rename_all(tolower)
```

```{r}
str(nyEconInd)
```

```{r}
nyEconInd$date <- mdy(nyEconInd$date)
```

```{r}
str(nyEconInd)
```

### e) Download US Census Data

```{r}
usCensus <- 
    read_csv("data/usCensus.csv") %>%  
    rename_all(tolower)
```


```{r}
head(usCensus)
```

```{r}
usCensus <-
  usCensus[-c(1)]
usCensus <-
  usCensus[-c(1),] 
usCensus <-
  usCensus[-c(52,53),]
usCensus <- arrange(usCensus, name)
```


```{r}
unique(usCensus$name)
```


```{r}
populationState <- usCensus %>% 
  filter(name!='District of Columbia') %>% 
  select(stateName = name, population = p001001) %>% 
  bind_cols(state = state.abb) %>% 
  bind_rows(usCensus %>% 
              filter(name=='District of Columbia') %>% 
              select(stateName=name, population=p001001) %>% 
              bind_cols(state='DC')) %>% 
  arrange(state) %>% 
  mutate(population = as.numeric(population))
```

```{r}
populationState
```

#### i) Add population column to tsLCOrg

```{r}

tsLC <- tsLCOrg %>% 
  inner_join(populationState, by='state')

```

```{r}
tsLC
```

```{r}
str(tsLC)
```

#### ii) Calculate Loans per capita
```{r}
tsLC <- tsLC %>% 
  mutate(loansPerCapita = totalloans/population)
```

```{r}
tsLC
```

#### iii)

```{r}
tsLC <- 
  tsLC %>% 
  left_join(nyEconInd, by=c("date"="date","state"="state"))
```

```{r}
filter(tsLC, state=="NY")
```

```{r}
tsLC <- as_tsibble(tsLC, key=state, index=date)
```


## 2) Exploratory Analysis

### a) Loans per Capita by State Plot 
```{r}
nintyPercentile = quantile(tsLC$population, 0.90)
tenPercentile = quantile(tsLC$population, 0.10)
```

```{r}
topTexBoxPlot <-
  tsLC %>% 
    filter(population >= nintyPercentile) %>% 
    ggplot(aes(x=factor(state), y=loansPerCapita)) + 
    ggtitle("loans per capita for Top Ten percentile States by population") +
    geom_boxplot()

topTexBoxPlot
ggplotly(topTexBoxPlot)
```

```{r}
bottomTexBoxPlot <-
  tsLC %>% 
    filter(population < tenPercentile) %>% 
    ggplot(aes(x=factor(state), y=loansPerCapita)) + 
    ggtitle("loans per capita for bottom Ten percentile States by population") +
    geom_boxplot()

bottomTexBoxPlot
ggplotly(bottomTexBoxPlot)
```

### b) Anomaly Plots

```{r}
library(anomalize)

anomalyLPCNY <- 
  tsLC %>%
  filter(state == "NY") %>% 
  time_decompose(loansPerCapita, method = "stl") %>%
  anomalize(remainder, method = "iqr") %>%
  plot_anomalies() +
  labs(title = "Anomaly detection for the loans per capita in NY") +
  xlab("Year") + ylab("Loans Per Capita ($)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")

anomalyLPCNY
```

```{r}
anomalyLPCMA <- 
  tsLC %>%
  filter(state == "MA") %>% 
  time_decompose(loansPerCapita, method = "stl") %>%
  anomalize(remainder, method = "iqr") %>%
  plot_anomalies() +
  labs(title = "Anomaly detection for the loans per capita in MA") +
  xlab("Year") + ylab("Loans Per Capita ($)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")

anomalyLPCMA
```

```{r}
anomalyLPCCO <- 
  tsLC %>%
  filter(state == "CO") %>% 
  time_decompose(loansPerCapita, method = "stl") %>%
  anomalize(remainder, method = "iqr") %>%
  plot_anomalies() +
  labs(title = "Anomaly detection for the loans per capita in CO") +
  xlab("Year") + ylab("Loans Per Capita ($)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")

anomalyLPCCO
```

### c) Apply STL Decomposition
#### i) For issues loans

```{r}
tsLC <- 
  tsLC %>% 
  mutate ( month = yearmonth(date)) 
```

```{r}
tsLC
```


```{r}

plotLPCNYDecomposed <- 
  tsLC %>%
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>% 
  model(STL(loansPerCapita ~ trend(window=10) + season(window='periodic'), robust = TRUE)) %>% 
  components() %>%
  autoplot() +
  xlab("Year") + ylab("Loans Per Capita ($)") +
  ggtitle("Seasonal and Trend decomposition using Loess (STL decomposition)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")
ggplotly(plotLPCNYDecomposed)
plotLPCNYDecomposed

```

### d)
```{r}

# Seasonal Plot
plotLPCNYSeason <-
  tsLC %>%
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>% 
  gg_season(loansPerCapita, labels = "both") +
  xlab("Month") + ylab("Loans Per Capita ($)") +
  ggtitle("Loans Per Capita in NY")
plotLPCNYSeason
```

```{r}
# Seasonal Sub Series Plot

plotLCPNYSub <- 
  tsLC %>% 
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>% 
  gg_subseries(loansPerCapita) +
  ylab("Loans Per Capita") +
  xlab("Year (monthly data)") +
  ggtitle("Seasonal subseries plot: Loans per Capita in NY")
plotLCPNYSub
```

#### Last 3 Years

```{r}

# Seasonal Plot
plotLPCNYSeason <-
  tsLC %>%
  filter(state=="NY" & year(date) >= 2014) %>% 
  as_tsibble (index = month) %>% 
  gg_season(loansPerCapita, labels = "both") +
  xlab("Month") + ylab("Loans Per Capita ($)") +
  ggtitle("Loans Per Capita in NY")
plotLPCNYSeason
```

```{r}
# Seasonal Sub Series Plot

plotLCPNYSub <- 
  tsLC %>% 
  filter(state=="NY" & year(date) >= 2014) %>% 
  as_tsibble (index = month) %>% 
  gg_subseries(loansPerCapita) +
  ylab("Loans Per Capita") +
  xlab("Year (monthly data)") +
  ggtitle("Seasonal subseries plot: Loans per Capita in NY")
plotLCPNYSub

```

### e) Autocorrelation Plots

```{r}
plotLPCNYACF <-
  tsLC %>% 
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>%
  ACF(loansPerCapita, lag_max = 48) %>%
  autoplot() +  ggtitle("Autocorrelation function (ACF) plot for loans per capita in NY")
plotLPCNYACF

```

```{r}
plotLPCNYPCF <-
  tsLC %>% 
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>%
  PACF(loansPerCapita, lag_max = 48) %>%
  autoplot() +  ggtitle("Partial Autocorrelation function (PACF) plot for Loans Per Capita in NY")
plotLPCNYPCF

```

### f) Lag Plot
```{r}
plotLPCNYLag <-
  tsLC %>%
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>%
  gg_lag(loansPerCapita, geom='point', lags = c(1,5,10,15,20,25)) +
  xlab(NULL) + ylab(NULL) +
  ggtitle("Lag plots for the Loans per Capita in NY")
ggplotly(plotLPCNYLag)
plotLPCNYLag
```

### g) Plots of loans per capita in NY over time

```{r}

plotLPCNY <-
  tsLC %>% 
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>%
  select(month, loansPerCapita) %>% 
  autoplot() +
  xlab("Year (monthly data)") + ylab("Loans Per Capita") +
  ggtitle("NY Loans Per Capita Data")
ggplotly(plotLPCNY)

plotLPCNY
```

```{r}
library("forecast")
LPCMovAvg <- data.frame(movingAvg =
  tsLC %>%
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>%
  select(month,loansPerCapita) %>% 
  ma(order=5))

plotMovAvg <-
  tsLC %>%
  filter(state=="NY") %>% 
  as_tsibble (index = month) %>%
  select(month,loansPerCapita) %>% 
  bind_cols(LPCMovAvg) %>% 
  ggplot()+
  geom_line(aes(y=movingAvg, x=month),color="blue") + 
  geom_line(aes(y =loansPerCapita, x = month), color = "red")

plotMovAvg
ggplotly(plotMovAvg)


```

## 3) Modeliing the loans issued in NY

### a) Seasonal naive and drift forecast

```{r}
#Creating a new dataframe for NY data with months as index
tsLPCNY <- 
  tsLC %>% 
  filter(state=="NY") %>% 
  as_tsibble (index = month)
  
```


```{r}
plotLPCNYNaive <-
  tsLPCNY %>%
  model(SNAIVE(loansPerCapita)) %>%
  forecast(h = "5 years") %>%
  autoplot(tsLPCNY, colour = "#769ECB") + #level = NULL,
  geom_line(linetype = 'dashed', colour = '#000000') +
  xlab("Year (monthly data)") + ylab("Loans Per Capita ($)") +
  ggtitle("NY Loans per capita data")
plotLPCNYNaive
```

```{r}
plotLPCNYDrift <-
  tsLPCNY %>%
  model(RW(loansPerCapita ~ drift())) %>%
  forecast(h = "5 years") %>%
  autoplot(tsLPCNY, colour = "#769ECB") + #level = NULL,
  geom_line(linetype = 'dashed', colour = '#000000') +
  xlab("Year (monthly data)") + ylab("Loans Per Capita ($)") +
  ggtitle("NY Loans per capita data")
plotLPCNYDrift
```

### b) Time Series Regression

```{r}
fitLPCNY <-
  tsLPCNY %>%
  model(TSLM(loansPerCapita ~ trend() + season() + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits))
report(fitLPCNY)
```


```{r}
skim(as_tibble(tsLPCNY))
```

### c) 

```{r}
# Plot of fitted values for above model
fitLPCNYFitted <-
  augment(fitLPCNY) %>%
  ggplot(aes(x = month)) +
  geom_line(aes(y = loansPerCapita, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  xlab("Year") + ylab("Loans Per Capita") +
  ggtitle("Loans Per Capita in NY") +
  scale_x_date(date_breaks = "years" , date_labels = "%y") +
  guides(colour=guide_legend(title=NULL))
ggplotly(fitLPCNYFitted)
```

```{r}
# Model without time components trend and seasonality
fitLPCNY_NoTrendSeason <-
  tsLPCNY %>%
  model(TSLM(loansPerCapita ~ avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits))
report(fitLPCNY_NoTrendSeason)
```

```{r}
# Plot of fitted values for above model
fitLPCNY_NoTrendSeasonFitted <-
  augment(fitLPCNY_NoTrendSeason) %>%
  ggplot(aes(x = month)) +
  geom_line(aes(y = loansPerCapita, colour = "Data")) +
  geom_line(aes(y = .fitted, colour = "Fitted")) +
  xlab("Year") + ylab("Loans Per Capita") +
  ggtitle("Loans Per Capita in NY") +
  scale_x_date(date_breaks = "years" , date_labels = "%y") +
  guides(colour=guide_legend(title=NULL))
ggplotly(fitLPCNY_NoTrendSeasonFitted)
```

### d)


```{r}
plotLPCNYPredicted_2014 <-
  tsLPCNY %>%
  filter(year(date) <= '2014') %>%
  model(TSLM(loansPerCapita ~ trend() + season() + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx +nysnapbenefits)) %>%
  forecast(new_data = tsLPCNY %>% filter(year(date) > '2014')) %>%
  autoplot(tsLPCNY, colour = "#960A0A") + #level = NULL,
  geom_line(colour = '#535353') +
  xlab("Year (monthly data)") + ylab("Loans per capita") +
  ggtitle("NY Loans per capita data")
plotLPCNYPredicted_2014
```

```{r}
plotLPCNYPredicted_2015 <-
  tsLPCNY %>%
  filter(year(date) <= '2015') %>%
  model(TSLM(loansPerCapita ~ trend() + season() + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx +nysnapbenefits)) %>%
  forecast(new_data = tsLPCNY %>% filter(year(date) > '2015')) %>%
  autoplot(tsLPCNY, colour = "#960A0A") + #level = NULL,
  geom_line(colour = '#535353') +
  xlab("Year (monthly data)") + ylab("Loans per capita") +
  ggtitle("NY Loans per capita data")
plotLPCNYPredicted_2015
```

### e) residual diagnostics
```{r}
fitLPCNY %>% gg_tsresiduals()
```

### f) ARIMA Model
```{r}
fitLPCNYARIMA <-
  tsLPCNY %>%
  model(fitArima = ARIMA(loansPerCapita ~ PDQ(0,0,0) + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits, #pdq(3,0,0) + 
              stepwise = FALSE, approximation = FALSE)
  )
report(fitLPCNYARIMA)
```

```{r}
# p-value for avg interest rate
2*pt(-abs(0.0209/0.0340), nrow(tsLPCNY)-15)
# p-value for avg term
2*pt(-abs(-0.0126/0.0127), nrow(tsLPCNY)-15)
# p-value for avg verification status
2*pt(-abs(-0.0779/0.1719), nrow(tsLPCNY)-15)
# p-value for avg home ownership
2*pt(-abs(-0.0003/0.4238), nrow(tsLPCNY)-15)
# p-value for NY CPI
2*pt(-abs(-0.0046/0.0024), nrow(tsLPCNY)-15)
# p-value for NY Unemployment
2*pt(-abs(-0.3322/0.0629), nrow(tsLPCNY)-15)
# p-value for NY Condo Price Index
2*pt(-abs(0.0091/0.0048), nrow(tsLPCNY)-15)

```

### g) 
```{r}
tsLPCNY %>%
  features(loansPerCapita, unitroot_kpss)
```

```{r}
tsLPCNY %>%
  features(loansPerCapita, unitroot_nsdiffs)
```


```{r}
tsLPCNY %>% 
  features(loansPerCapita, unitroot_ndiffs)
```

### h)

```{r}
fitLPCNYARIMADiff <-
  tsLPCNY %>%
  model(fitArima = ARIMA(loansPerCapita ~ PDQ(0,0,0) + pdq(,1,) + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits, #pdq(3,0,0) + 
              stepwise = FALSE, approximation = FALSE)
  )
report(fitLPCNYARIMADiff)
```

## 4) Predictive modelling of the loans issues in NY
### a)

```{r}
tsLPCNY
```

```{r}
set.seed(333)
tsLPCNYTrain <- tsLPCNY %>% filter(date < '2016-03-01')
tsLPCNYTest <- tsLPCNY %>% filter(date >= '2016-03-01')
```


```{r}
tsLPCNYFitAll <-
  tsLPCNYTrain %>%
  model(
    model1TimeTrendAndSeason = TSLM(loansPerCapita ~ trend() + season()),
    model2WithVariables = TSLM(loansPerCapita ~ trend() + season() + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits),
    model3ArimaGrid = ARIMA(loansPerCapita ~ PDQ(0,0,0), stepwise = FALSE, approximation = FALSE),
    model4ArimaFrom3f = ARIMA(loansPerCapita ~ PDQ(0,0,0) + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits, stepwise = FALSE, approximation = FALSE)
  )

tsLPCNYPredictAll <-
  tsLPCNYFitAll %>%
  forecast(new_data = tsLPCNYTest)

accuracy(tsLPCNYPredictAll, tsLPCNYTest)
```

### b)
```{r}
set.seed(333)
tsLPCNYTrain_new <- tsLPCNY %>% filter(date < '2016-04-01')
tsLPCNYTest_new <- tsLPCNY %>% filter(date >= '2016-04-01')
```


```{r}
tsLPCNYFitAll <-
  tsLPCNYTrain_new %>%
  model(
    model1TimeTrendAndSeason = TSLM(loansPerCapita ~ trend() + season()),
    model2WithVariables = TSLM(loansPerCapita ~ trend() + season() + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits),
    model3ArimaGrid = ARIMA(loansPerCapita ~ PDQ(0,0,0), stepwise = FALSE, approximation = FALSE),
    model4ArimaFrom3f = ARIMA(loansPerCapita ~ PDQ(0,0,0) + avgintrate + avgterm + avgannualinc + avgverifstatus + avghomeowner + avgrevolbal + nycpi +nyunemployment + nycondopriceidx + nysnapbenefits, stepwise = FALSE, approximation = FALSE)
  )

tsLPCNYPredictAll <-
  tsLPCNYFitAll %>%
  forecast(new_data = tsLPCNYTest_new)

accuracy(tsLPCNYPredictAll, tsLPCNYTest_new)
```

# Part II
## 1) Preparation and exploration
### a) Load the data
```{r}
tsRetail <-
  read_csv("data/retailSales.csv") %>%  
  rename_all(tolower)
```

```{r}

tsRetail$date <- mdy(tsRetail$date)
```

### b)
```{r}

tsRetail <- as_tsibble(tsRetail, index=date)
```

### c)
#### i) Whole Data
```{r}
plotRetail <-
  tsRetail %>% 
  autoplot() +
  xlab("Year (monthly data)") + ylab("Retail Sales (in million dollars)") +
  ggtitle("U.S. Retails Sales data")
ggplotly(plotRetail)
plotRetail
```

#### ii) Subset of data for points after year 2010
```{r}
plotRetail_2010 <-
  tsRetail %>% 
  filter(year(date)>=2010) %>% 
  autoplot() +
  xlab("Year (monthly data)") + ylab("Retail Sales (in million dollars)") +
  ggtitle("U.S. Retails Sales data")
ggplotly(plotRetail_2010)
plotRetail_2010
```

## 2) Understanding the time series

### a) Seasonal and Seasonal Subseries plots

```{r}
tsRetail <- 
  tsRetail %>% 
  mutate ( month = yearmonth(date)) 
```

```{r}
# Seasonal Plot
plotRetailSeason <-
  tsRetail %>%
  filter( year(date) >= 2015) %>% 
  as_tsibble (index = month) %>% 
  gg_season(sales, labels = "both") +
  xlab("Month") + ylab("Retail Sales") +
  ggtitle("U.S. Retail Sales data")
plotRetailSeason
```

```{r}
plotRetailSub <- 
  tsRetail %>% 
  filter(year(date) >= 2015) %>% 
  as_tsibble (index = month) %>% 
  gg_subseries(sales) +
  ylab("Retail Sales") +
  xlab("Year (monthly data)") +
  ggtitle("Seasonal subseries plot: U.S. Retail sales data")
plotRetailSub
```

### b) STL decomposition plots
#### i) Full Data
```{r}

plotRetailDecomposed <- 
  tsRetail %>%
  as_tsibble (index = month) %>% 
  model(STL(sales ~ trend(window=10) + season(window='periodic'), robust = TRUE)) %>% 
  components() %>%
  autoplot() +
  xlab("Year") + ylab("Retail Sales") +
  ggtitle("Seasonal and Trend decomposition using Loess (STL decomposition)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")
ggplotly(plotRetailDecomposed)
plotRetailDecomposed
```

#### ii) Subset between 2005 and 2015
```{r}
plotRetailDecomposed_subset <- 
  tsRetail %>%
  filter(year(date)>=2005 & year(date) <=2015) %>% 
  as_tsibble (index = month) %>% 
  model(STL(sales ~ trend(window=10) + season(window='periodic'), robust = TRUE)) %>% 
  components() %>%
  autoplot() +
  xlab("Year") + ylab("Retail Sales") +
  ggtitle("Seasonal and Trend decomposition using Loess (STL decomposition)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")
ggplotly(plotRetailDecomposed_subset)
plotRetailDecomposed_subset
```

### c) ACF and PCF
```{r}
plotRetailACF <-
  tsRetail %>% 
  as_tsibble (index = month) %>%
  ACF(sales, lag_max = 48) %>%
  autoplot() +  ggtitle("Autocorrelation function (ACF) plot for Retail Sales data")
plotRetailACF

```

```{r}
plotRetailPCF <-
  tsRetail %>% 
  as_tsibble (index = month) %>%
  PACF(sales, lag_max = 48) %>%
  autoplot() +  ggtitle("Partial Autocorrelation function (PACF) plot for Retail Sales data")
plotRetailPCF
```

### d)
```{r}
plotRetailSeasonallyAdjusted <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  autoplot(sales, color='#A9A9B0') +
  autolayer(components(tsRetail %>% as_tsibble (index = month) %>%model(STL(sales))), season_adjust, color='#1490D4') +
  xlab("Year (monthly data)") + ylab("Retail Sales") +
  ggtitle("U.S. retail Sales data")
ggplotly(plotRetailSeasonallyAdjusted)
plotRetailSeasonallyAdjusted
```

### e)
```{r}

RetailMovAvg <- data.frame(movingAvg =
  tsRetail %>%
  as_tsibble (index = month) %>%
  select(sales) %>% 
  ma(order=2))

plotRetailMovAvg <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  select(month,sales) %>% 
  bind_cols(RetailMovAvg) %>% 
  ggplot()+
  geom_line(aes(y =sales, x = month), color = "#A9A9B0") +
  geom_line(aes(y=movingAvg, x=month),color="#1490D4")

plotRetailMovAvg
ggplotly(plotRetailMovAvg)
```

```{r}
# Changes to achieve a plot similar to plot in 2(d)
RetailMovAvg <- data.frame(movingAvg =
  tsRetail %>%
  as_tsibble (index = month) %>%
  select(sales) %>% 
  ma(order=7))

plotMovAvg <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  select(month,sales) %>% 
  bind_cols(RetailMovAvg) %>% 
  ggplot()+
  geom_line(aes(y =sales, x = month), color = "#A9A9B0") +
  geom_line(aes(y=movingAvg, x=month),color="#1490D4") 
  

plotMovAvg
ggplotly(plotMovAvg)
```

## 3) Modelling and analysis of the time series
### a) Time Series Regression

```{r}
fitRetail <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  model(TSLM(sales ~ trend() + season()))
report(fitRetail)
```

```{r}
fitRetail %>% gg_tsresiduals()
```


### b) ARIMA Model

```{r}
fitRetailARIMADiff <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  model(fitArima = ARIMA(sales, stepwise = FALSE, approximation = FALSE)
  )
report(fitRetailARIMADiff)
```

```{r}
fitRetailARIMADiff %>% gg_tsresiduals()
```

### c)
```{r}
tsRetail %>% 
  as_tsibble (index = month) %>%
  features(sales, unitroot_ndiffs)
```

```{r}
tsRetail %>%
  as_tsibble (index = month) %>%
  features(sales, unitroot_nsdiffs)
```

```{r}
tsRetail %>% 
  as_tsibble (index = month) %>%
  features(difference(difference(sales,12)), unitroot_kpss)
```

```{r}
plotRetailACF <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  PACF(sales, lag_max = 48) %>%
  autoplot() +  ggtitle("PACF plot for the retail sales data")
plotRetailACF
```

```{r}
plotRetailACF <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  PACF(difference(difference(sales),12), lag_max = 48) %>%
  autoplot() +  ggtitle("PACF plot for the retail sales data")
plotRetailACF
```


```{r}
plotRetailACF <-
  tsRetail %>%
  as_tsibble (index = month) %>%
  PACF(difference(difference(sales,12)), lag_max = 48) %>%
  autoplot() +  ggtitle("PACF plot for the retail sales data")
plotRetailACF
```

### d)
```{r}
set.seed(333)
tsRetailTrain <- tsRetail %>% as_tsibble (index = month) %>%filter(year(date) < '2011')
tsRetailTest <- tsRetail %>% as_tsibble (index = month) %>%filter(year(date) >= '2011')
```

```{r}
tsRetailFitAll10Yr <-
  tsRetailTrain %>%
  model(
    model1TimeTrendAndSeason = TSLM(sales ~ trend() + season()),
    model2ArimaGrid = ARIMA(sales,stepwise = FALSE, approximation = FALSE)
  )

tsRetailPredictAll10Yr <-
  tsRetailFitAll10Yr %>%
  forecast(new_data = tsRetailTest)

accuracy(tsRetailPredictAll10Yr, tsRetailTest)

```

### e)

```{r}
set.seed(333)
tsRetailTrain_2016 <- tsRetail %>% as_tsibble (index = month) %>%filter(year(date) < '2016')
tsRetailTest_2016 <- tsRetail %>% as_tsibble (index = month) %>%filter(year(date) >= '2016')
```


```{r}
tsRetailFitAll5Yr <-
  tsRetailTrain_2016 %>%
  model(
    model1TimeTrendAndSeason = TSLM(sales ~ trend() + season()),
    model2ArimaGrid = ARIMA(sales,stepwise = FALSE, approximation = FALSE)
  )

tsRetailPredictAll5Yr <-
  tsRetailFitAll5Yr %>%
  forecast(new_data = tsRetailTest_2016)

accuracy(tsRetailPredictAll5Yr, tsRetailTest_2016)
```

## 4) Checking for anomalies and reporting the results
### a) 
```{r}
library(anomalize)

anomalyRetail <- 
  tsRetail %>%
  time_decompose(sales, method = "stl") %>%
  anomalize(remainder, method = "gesd") %>%
  plot_anomalies() +
  labs(title = "Anomaly detection for Retail Sales data") +
  xlab("Year") + ylab("Retail Sales ($ million)") +
  scale_x_date(date_breaks = "years" , date_labels = "%y")

anomalyRetail
```

```{r}
anomalyList <-
  tsRetail %>%
  time_decompose(sales, method = "stl") %>%
  anomalize(remainder, method = "gesd")

as_tibble(anomalyList) %>% 
  filter(anomaly == 'Yes')
```

### b)
```{r}
timeTrendPrediction <-
  as_tsibble(tsRetailPredictAll10Yr) %>% 
  filter(.model=="model1TimeTrendAndSeason") %>% 
  mutate(timeTrendSales = sales) %>% 
  select(date, timeTrendSales) %>% 
  as_tsibble(index = date)

arimaPrediction <-
  as_tsibble(tsRetailPredictAll10Yr) %>% 
  filter(.model=="model2ArimaGrid") %>% 
  mutate(arimaSales = sales) %>% 
  select(date, arimaSales) %>% 
  as_tsibble(index = date)
```

```{r}
tsPlot <- tsRetail %>% 
  filter(year(date)>=2010)  %>% 
  left_join(arimaPrediction, by='date') %>% 
  left_join(timeTrendPrediction, by='date')
```

```{r}
plotAll <-
  tsPlot %>%
  #as_tsibble (index = month.x) %>%
  ggplot()+
  geom_line(aes(y=sales, x=month.x, colour= "Actual Data")) + 
  geom_line(aes(y =timeTrendSales, x = month.x, colour = "Forecast Time Trend Model")) +
  geom_line(aes(y =arimaSales, x = month.x, colour = "Forecast Arima Model"))

plotAll
ggplotly(plotAll)
```


```{r}
timeTrendPrediction <-
  as_tsibble(tsRetailPredictAll5Yr) %>% 
  filter(.model=="model1TimeTrendAndSeason") %>% 
  mutate(timeTrendSales = sales) %>% 
  select(date, timeTrendSales) %>% 
  as_tsibble(index = date)

arimaPrediction <-
  as_tsibble(tsRetailPredictAll5Yr) %>% 
  filter(.model=="model2ArimaGrid") %>% 
  mutate(arimaSales = sales) %>% 
  select(date, arimaSales) %>% 
  as_tsibble(index = date)
```

```{r}
tsPlot1 <- tsRetail %>% 
  filter(year(date)>=2010)  %>% 
  left_join(arimaPrediction, by='date') %>% 
  left_join(timeTrendPrediction, by='date')
```

```{r}
plotAll1 <-
  tsPlot1 %>%
  #as_tsibble (index = month.x) %>%
  ggplot()+
  geom_line(aes(y=sales, x=month.x, colour= "Actual Data")) + 
  geom_line(aes(y =timeTrendSales, x = month.x, colour = "Forecast Time Trend Model")) +
  geom_line(aes(y =arimaSales, x = month.x, colour = "Forecast Arima Model"))

plotAll1
ggplotly(plotAll1)

```

```{r}
tsRetail
```

## Bonus Question:
### 1)

```{r}
tsEconUs <-
  read_csv("data/usEcon.csv") %>%  
  rename_all(tolower)
```

```{r}
tsEconUs$date <- mdy(tsEconUs$date)
```

```{r}
tsEconRetail <- 
  tsRetail %>% 
  inner_join(tsEconUs, by="date")
```

```{r}
tsEconRetail$inflation <- as.numeric(sub("%", "", tsEconRetail$inflation))
```


```{r}
tsEconRetail
```


```{r}
set.seed(333)
tsEconRetailTrain <- tsEconRetail %>% as_tsibble (index = month) %>%filter(year(date) < '2016')
tsEconRetailTest <- tsEconRetail %>% as_tsibble (index = month) %>%filter(year(date) >= '2016')
```

```{r}
tsEconRetailFitAll <-
  tsEconRetailTrain %>%
  model(
    model1TimeTrendAndSeason = TSLM(sales ~ trend() + season() + inflation + cpi + income + unemployment),
    model2ArimaGrid = ARIMA(sales ~ inflation + cpi + income + unemployment,stepwise = FALSE, approximation = FALSE)
  )

tsEconRetailPredictAll <-
  tsEconRetailFitAll %>%
  forecast(new_data = tsEconRetailTest)

accuracy(tsEconRetailPredictAll, tsEconRetailTest)

```



